---
title: Major employment data
---

We extract a number of of the priority sites from Data Axle data; that extraction is done in this notebook.

```{r}
library(tidyverse)
library(sf)
library(RcppTOML)

config = parseTOML(here::here("config.toml"))
```

```{r}
data = read_csv(file.path(config$dataaxle_path, "NC_bus_2022.csv"), guess_max=1000000, col_types=cols("NAICS6"=col_character())) %>%
    filter(`COUNTY CODE` == 119)
```

## Major employment

Filter to just >100 employees in Mecklenburg County. Note: for some reason this only turns up 832 records, vs. 984 in the search interface. Ask Phil McDaniel about this.

```{r}
major_emp = data %>%
    filter(
        !is.na(`EMPLOYEE SIZE (5) - LOCATION`) &
        `EMPLOYEE SIZE (5) - LOCATION` >= 100
    ) %>%
    mutate(LATITUDE=as.numeric(LATITUDE)) %>%
    select(COMPANY, `EMPLOYEE SIZE (5) - LOCATION`, LATITUDE, LONGITUDE) %>%
    st_as_sf(coords=c("LONGITUDE", "LATITUDE"), crs=4326)

write_sf(major_emp, file.path(config$dataaxle_path, "mecklenburg_large_employers.gpkg"))
```

## Hotels

We include all "traveler accomodation" codes.

```{r}
hotels = data %>%
    # 7211 - traveler accomodation
    # 7213 - rooming and boarding houses, dormitories, and workers' camps - I looked at some of these and
    # they are mostly student/corporate housing
    filter(str_starts(NAICS6, "7211")) %>%
        mutate(LATITUDE=as.numeric(LATITUDE)) %>%
        select(COMPANY, NAICS6, LATITUDE, LONGITUDE) %>%
        st_as_sf(coords=c("LONGITUDE", "LATITUDE"), crs=4326) %>%
        write_sf(file.path(config$dataaxle_path, "mecklenburg_hotels.gpkg"))
```

## Retail

In the official methodology, this is defined as a "retail center" which usually means a strip mall. We don't have a reliable way of identifying those from the Data Axle data, so we just use all retail and bar/restaurant locations, but weight them down by a factor of 20%—estimating that the typical strip mall has five stores. This also means we are not discounting downtown retail (note: now that I think about it the typical strip mall probably has more than 5 stores). We are using the following NAICS codes:

    - 44, 45 except 445132 Vending machine operators
    - 7224 Drinking places (alcoholic beverages)
    - 7225 Restaurants and other eating places

TODO any way to separate out headquarters/distribution centers?

```{r}
data %>%
    filter(
        ((str_starts(NAICS6, "44") | str_starts(NAICS6, "45")) & NAICS6 != "445132") |
        str_starts(NAICS6, "7224") |
        str_starts(NAICS6, "7225")
    ) %>%
    mutate(LATITUDE=as.numeric(LATITUDE)) %>%
    select(COMPANY, NAICS6, LATITUDE, LONGITUDE) %>%
    st_as_sf(coords=c("LONGITUDE", "LATITUDE"), crs=4326) %>%
    write_sf(file.path(config$dataaxle_path, "mecklenburg_retail.gpkg"))
```

## Community/religious/conference

813110 is church or religious org - unfortunately some of these appear to just be people's houses, but I don't see a good way to address that

Community/civic centers come from https://parkandrec.mecknc.gov/Places-to-Visit/Rec-Centers


```{r}
data %>%
    filter(NAICS6 == "813110") %>%
    mutate(LATITUDE=as.numeric(LATITUDE)) %>%
    select(COMPANY, NAICS6, `EMPLOYEE SIZE (5) - LOCATION`, LATITUDE, LONGITUDE) %>%
    st_as_sf(coords=c("LONGITUDE", "LATITUDE"), crs=4326) %>%
    write_sf(file.path(config$dataaxle_path, "mecklenburg_church.gpkg"))
```

