---
title: Colleges and universities
---

We use the [IPEDS shapefile](https://nces.ed.gov/programs/edge/geographic/schoollocations) and extract schools in Mecklenburg County.

```{r}
library(sf)
library(tidyverse)
library(tigris)
library(RcppTOML)

config = parseTOML(here::here("config.toml"))
```

```{r}
unis = read_sf(file.path(config$data_path, "data", "destinations", "EDGE_GEOCODE_POSTSECSCH_2223.shp")) %>%
    filter(CNTY == "37119") # just mecklenburg county
```

Some of these are not actually colleges/universities (e.g. the beauty schools). Merge with [IPEDS directory information](https://nces.ed.gov/ipeds/datacenter/DataFiles.aspx?gotoReportId=7&fromIpeds=true&sid=e273376e-ffab-4855-9830-616bd2f03635&rtid=7) and remove any school that does not grant at least associates degrees.

```{r}
ipeds = read_sf(file.path(config$data_path, "data", "destinations", "hd2022.csv"))
```

The IPEDS data is for entire institutions while the geocodes are for individual campuses, which have a two digit code appended.

```{r}
unis = unis %>%
    mutate(main_unitid = str_sub(UNITID, 1, 6)) %>%
    left_join(ipeds, by=c("main_unitid"="UNITID"))

stopifnot(!any(is.na(unis$ICLEVEL)))
```


```{r}
unis = unis %>%
    filter(ICLEVEL %in% c("1", "2"))
```

```{r}
unis %>% write_sf(file.path(config$data_path, "data", "destinations", "colleges.gpkg"))
```